Measuring matter effects in Neutron star - Black hole binaries with Advanced LIGO

### 2015-11-11 Paper/paper.tex
add figures showing the SNR levels above which we can make meaningful statements on NS Lambda

% Macros for text notes and comments
\newcommand{\Note}[1]{\textcolor{blue}{\textbf{[#1]}}}

@@

\begin{figure*}[!t]
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the difference that we consider the systematic bias in the dimensionless
mass-ratio $\eta$ here. In the bottom row we show control runs where both
signal and template waveforms lack matter effects.}
\label{fig:TNT_EtaBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
  with the difference that we consider the confidence intervals in the recovery
  of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
  Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
  confidence interval is shown here.}
\label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
\caption{The top two rows in this figure are
similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the differences that we consider the systematic bias in the dimensionless
spin on the BH here, and that the differences shown are absolute and not
fractions of the injected spin. In the bottom row, we show a control run where
both injected and template waveforms correspond to binary black holes.}
\label{fig:TNT_BHspinBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
\caption{The top two rows in the figure are similar to
Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
with the difference that we consider the confidence intervals in the recovery
of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
width of the confidence interval without normalizing with the injected spin.
The bottom row show a control run where matter effects are ignored both 
in the injected signal and the template waveforms.}
\label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
\caption{This figure is similar to the bottom row of
Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
consider the NS tidal parameter $\Lambda$ here.}
\label{fig:TT_NSLambdaBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
\caption{This figure is similar to the bottom row of
Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
we consider the NS tidal parameter $\Lambda$ here.}
\label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
\end{figure*}
% 
% 
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.

@@

Could have subsections for aLIGO and ET ?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What do we gain by using templates that include NS matter effects?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Having shown in the previous section that we begin to care for NS tidal effects
for signals in the XXX corner of the paraemter space, with SNRs above YY, here
we investigate the effect of the improvement in the accuracy of the recovered
parameters when tidal effects are included in the templates.

\begin{enumerate}
\item What is the reduction in the bias of the maximum likelihood parameters
when using tidal templates?\newline
``Plot the ratio of the bias between N templates and T templates (against T
signals), as a function of SNR.''
\item What is the reduction in the uncertainty in binary mass and spin, if
any?\newline
``show how the the $90\%$ confidence intervals shrink, as a function of SNR, 
when we go from using N templates to T templates.''
\end{enumerate}

\red{Could have subsections for aLIGO and ET ?}

>>>>

% Macros for text notes and comments
\newcommand{\Note}[1]{\textcolor{blue}{\textbf{[#1]}}}
\newcommand{\lambdans}{\Lambda_\mathrm{NS}}
\newcommand{\chibh}{\chi_\mathrm{BH}}
\newcommand{\mbh}{m_\mathrm{BH}}
\newcommand{\mns}{m_\mathrm{NS}}

@@

\begin{figure*}[!t]
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaBias_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the difference that we consider the systematic bias in the dimensionless
mass-ratio $\eta$ here. In the bottom row we show control runs where both
signal and template waveforms lack matter effects.}
\label{fig:TNT_EtaBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_EtaCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_EtaCIWidth90_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_EtaCIWidth90_vs_SNR_q23.pdf} 
\caption{This figure (top two rows) is similar to Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
  with the difference that we consider the confidence intervals in the recovery
  of the dimensionless mass-ratio $\eta$ here. The bottom row here is similar to 
  Fig.~\ref{fig:TNT_EtaBias_vs_SNR_q23} with the difference that width of the $90\%$
  confidence interval is shown here.}
\label{fig:TNT_EtaCIWidth90_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinBias_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinBias_vs_SNR_q23.pdf}\\%\quad
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinBias_vs_SNR_q23.pdf}
\caption{The top two rows in this figure are
similar to Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23},
with the differences that we consider the systematic bias in the dimensionless
spin on the BH here, and that the differences shown are absolute and not
fractions of the injected spin. In the bottom row, we show a control run where
both injected and template waveforms correspond to binary black holes.}
\label{fig:TNT_BHspinBias_vs_SNR_q23}
\end{figure*}
% 
\begin{figure*}
\centering    
\textbf{BH-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/NN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; BH-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TN_BHspinCIWidth90_vs_SNR_q23.pdf}\\ 
\textbf{NS-BH Injection; NS-BH Templates}
\includegraphics[width=1.7\columnwidth]{plots/TT_BHspinCIWidth90_vs_SNR_q23.pdf}
\caption{The top two rows in the figure are similar to
Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23},
with the difference that we consider the confidence intervals in the recovery
of the dimensionless spin on the BH $\chi_\mathrm{BH}$ here.
As in Fig.~\ref{fig:TNT_BHspinBias_vs_SNR_q23}, we show here the absolute 
width of the confidence interval without normalizing with the injected spin.
The bottom row show a control run where matter effects are ignored both 
in the injected signal and the template waveforms.}
\label{fig:TNT_BHspinCIWidth90_vs_SNR_q23}
\end{figure*}
% 
% 
Advanced LIGO searches and parameter estimation efforts aimed at binaries 
containing NS and stellar-mass BHs are poised to use waveform models that do not
include the effects of the NS tidal deformability~\cite{Canton:2014ena}.
While this is not expected to
be the dominant source of error at low SNRs, it will likely introduce a
systematic bias in the physical parameters that we recover from GW observations.
In this section, we study the impact of the same for binaries where the BH spins
are aligned to the orbital angular momentum.

@@

\prayush{Could have subsections for different frequency-optimized aLIGO noise curves ?}

\hspace{5mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What do we gain by using templates that include NS matter effects?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % 
% \begin{figure*}
% \centering
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaBias_vs_SNR_q23.pdf}
% \caption{This figure is similar to the bottom row of
% Fig.~\ref{fig:TNT_chirpMassBias_vs_SNR_q23}, with the difference that  we
% consider the NS tidal parameter $\Lambda$ here.}
% \label{fig:TT_NSLambdaBias_vs_SNR_q23}
% \end{figure*}
% % 
% \begin{figure*}
% \centering
% \textbf{NS-BH Injection; NS-BH Templates}
% \includegraphics[width=1.7\columnwidth]{plots/TT_NSLambdaCIWidth90_vs_SNR_q23.pdf}
% \caption{This figure is similar to the bottom row of
% Fig.~\ref{fig:TNT_chirpMassCIWidth90_vs_SNR_q23}, with the difference that  
% we consider the NS tidal parameter $\Lambda$ here.}
% \label{fig:TT_NSLambdaCIWidth90_vs_SNR_q23}
% \end{figure*}
% % 
% \begin{figure*}
% \centering    
% \includegraphics[width=2\columnwidth]{plots-TT/TT_LambdaBiases_CI90_0_Lambda_SNR.png}
% \caption{This figure shows.}
% \label{fig:TT_LambdaBiases_CI90_0_Lambda_SNR}
% \end{figure*}
% 
\begin{figure*}
\centering    
\includegraphics[width=2.1\columnwidth]{plots-TT/TT_LambdaCIWidths68_3_Lambda_SNR.png}\\
\includegraphics[width=2.1\columnwidth]{plots-TT/TT_LambdaCIWidths90_0_Lambda_SNR.png}
\caption{This figure shows.}
\label{fig:TT_LambdaCIWidths90_0_Lambda_SNR}
\end{figure*}
% 
\begin{figure*}
\centering    
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI68_3.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI68_3.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI68_3.png}\\
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda500_0_CI90_0.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda800_0_CI90_0.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_Lambda1000_0_CI90_0.png}
\caption{This figure shows.}
\label{fig:TT_SNRThresholdForLambdaMeasurement_BHspin_MassRatio_CI90_0}
\end{figure*}
% 
\begin{figure*}
\centering    
\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR30_0_CI68_3.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI68_3.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI68_3.png}\\
\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR50_0_CI90_0.png}
\includegraphics[width=.65\columnwidth]{plots-TT/TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass_SNR90_0_CI90_0.png}
\caption{This figure shows.}
\label{fig:TT_LambdaThresholdForLambdaMeasurement_BHspin_BHmass}
\end{figure*}

Having shown in the previous section that we begin to care for NS tidal effects
for signals in the XXX corner of the paraemter space, with SNRs above YY, here
we investigate the effect of the improvement in the accuracy of the recovered
parameters when tidal effects are included in the templates.

\begin{enumerate}
\item Where in the parameter space can we actually make a statement about 
$\Lambda_\mathrm{NS}$ ?
\item What is the reduction in the bias of the maximum likelihood parameters
when using tidal templates?\newline
``Plot the ratio of the bias between N templates and T templates (against T
signals), as a function of SNR.''
\item What is the reduction in the uncertainty in binary mass and spin, if
any?\newline
``show how the the $90\%$ confidence intervals shrink, as a function of SNR, 
when we go from using N templates to T templates.''
\end{enumerate}

\subsection{Constraining the tidal deformability of the neutron star}\label{s2:measuring_lambda}

With the enhanced sensitivity of aLIGO, we are likely to be able to measure or
constrain deviations in the observed GW signals, especially close to binary
merger, from their expected form under the point-particle approximation.
The tidal disruption of the NS close to merger, and outside of the 
inner-most circular orbit, will drastically change the morphology of the emitted
GW signal, the measurement of which will allow us to constrain the equation 
of state (EoS) of nuclear matter.
The earlier in the orbit the NS disrupts, the sharper is the drop in the emitted
GW signal for frequencies above the disruption frequency. In this section we
study the ability of aLIGO to constrain the 
$\lambdans\sim \left(R/M\right)_\mathrm{NS}^5$ parameter, which
the combination that parametrizes the deviation in the emitted GW signal.
Past studies~\cite{}
have studied a limited sample of systems with distinguishability criterion that
are typically applicable in the very high signal-to-noise ratio cases. 
In what follows, we will use stochastic parameter estimation algorithms to show
aLIGO's capability of extracting EoS information out from GW signals with 
more realistic (likely) SNRs.

As before we focus on binaries with mass-ratios $q=\{[2,3,4,5\}$, with the NS mass
fixed at $1.35M_\odot$. We allow BH spin to take the values 
$\chibh=\{-0.5,0,+0.5,+0.75\}$, and $\lambdans=\{500, 800, 1000\}$. For each of these
parameter combinations, we inject a tidal-corrected waveform using the 
Lackey et al~\cite{} model , and use MCMC with ensemble sampling to construct the 
posterior probability distribution for different binary parameters with (a) BHBH
templates, and (b) NSBH templates. We will discuss the results of (b) here.
